




```{r}
library(dplyr)
library(readxl)
library(tidyr)
library(moments)
```

```{r}
df <- readxl::read_excel("Zad_domowe_nr_1_2023_2024_KCh.xlsx")
```

```{r}
head(df,5)
```

```{r}
glimpse(df)
```

```{r}
df %>%
  summarize(
    srednia_masa = mean(Masa_ptaka),
    wariancja_masy = var(Masa_ptaka) * (nrow(df) - 1) * (1 / nrow(df)),
    odch_stand_masy = sd(Masa_ptaka) * (nrow(df) - 1) * (1 / nrow(df)),
    pierw_kwart_masy = quantile(Masa_ptaka, 0.25),
    mediana_masy = quantile(Masa_ptaka, 0.5), 
    trz_kwart_masy = quantile(Masa_ptaka, 0.75)
  )


```

```{r}
calc_skewness_coeff <- function(numbers) {
  
  xbar <- mean(numbers)
  n <- length(numbers)
  stand_dev <- sd(numbers) * (n - 1) / n
  num <- (1 / n) * sum((numbers - xbar) ^ 3)
  denom <- stand_dev ^ 3
  skew <- num / denom
  
  return(skew)
}
```

```{r}
quartile_coeff_of_dispersion <- function(numbers) {
  
  first.quart <- quantile(numbers, 0.25, names = FALSE)
  third.quart <- quantile(numbers, 0.75, names = FALSE)
  
  numerator = third.quart - first.quart
  denominator = third.quart + first.quart
  
  quartile.dispersion = numerator / denominator
  
  return(quartile.dispersion)
}
```

```{r}
quartile_coeff_of_dispersion(df$Masa_ptaka)
```

```{r}
?quantile
```

```{r}
```




```{r}

df %>%
  summarize(
    rozstep_miedzykwartylowy = IQR(Masa_ptaka), 
    odch_cwiartkowe = IQR(Masa_ptaka) / 2,
    skosnosc = skewness(Masa_ptaka),
    pozycyjna_skosnosc = quartile_coeff_of_dispersion(Masa_ptaka)
    
  )
```

```{r}
calc_skewness_coeff(df$Masa_ptaka)
```

```{r}
skewness(df$Masa_ptaka)
```

```{r}
boxplot(df$Masa_ptaka, range = 0)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
